---
title: "SPRIS Project 3"
author: "Jessica Lavery"
date: "3/31/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(tidyverse)
library(gtsummary)
library(lme4)
library(mice)

# vignette for mice package for longitudinal data
# https://www.gerkovink.com/miceVignettes/Multi_level/Multi_level_data.html

load(here::here("datasets.rdata"))

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

```{r}
# set theme for gtsummary
lst_theme_lavery <- list(
  "pkgwide-str:theme_name" = "Jessica Lavery",
  # round large pvalues to 2 places
  "pkgwide-fn:pvalue_fun" = function(x) gtsummary::style_pvalue(x, digits = 1),
  "pkgwide-fn:prependpvalue_fun" = function(x) gtsummary::style_pvalue(x, digits = 2, prepend_p = TRUE),
  # as_gt additional commands
  "as_gt-lst:addl_cmds" = list(gt = rlang::expr(gt::tab_options(table.font.size = "small", data_row.padding = gt::px(0)))),
  # flextable formatting
  "as_flex_table-lst:addl_cmds" = list(autofit = list(
    rlang::expr(flextable::font(fontname = "Arial", part = "all")),
    rlang::expr(flextable::fontsize(size = 11, part = "all"))
  )),
  # display a greek beta as header in tbl_regression
  "tbl_regression-str:coef_header" = rlang::expr(ifelse(exponentiate == TRUE, "exp(\U03B2)", "\U03B2"))
)
set_gtsummary_theme(lst_theme_lavery)
```


# Table 1

```{r}
data_expanded %>% 
  distinct(subject_id, tx, age, gender) %>% 
  select(-subject_id) %>% 
  tbl_summary(by = tx,
              label = list(age = "Age",
                           gender = "Gender")) %>% 
  add_overall() %>% 
  bold_labels() %>% 
  italicize_levels()
```

# Spaghetti plot for patient trajectories over time

Does not seem that a random slope is necessary?

```{r}
data_expanded %>% 
  mutate(measurement_time = case_when(
    day == 0 ~ 1,
    day == 5 ~ 2,
    day == 19 ~ 3,
    day == 90 ~ 4
  )) %>% 
  ggplot(aes(x = day, y = mem_comp)) +
  geom_point(aes(color = treatment_group, group = subject_id)) +
  geom_line(aes(color = treatment_group, group = subject_id)) +
  geom_smooth(color = "gray", size = 2, method = "loess", se = FALSE) +
  labs(x = "Day",
       color = "Treatment",
       y = "Cognitive Rating",
       caption = "Higher cognitive rating indicates better memory.") +
  facet_grid(.~treatment_group)
```

# Numeric summary of average cognitive scores over time, by treatment

```{r}
data_expanded %>% 
  tbl_summary(by = treatment_group,
              include = c(treatment_group, mem_comp))
```


# Distribution of missing data

This figure doesn't help a ton to identify patterns since only 1 var is missing data

```{r}
visdat::vis_miss(data_expanded)
```

From Mice package, but weird output? 

```{r}
md.pattern(data_expanded)
```

Does the distribution of missing data depend on the covariates?

Data more likely to be missing at later time points

Otherwise no significant differences in the distribution of missingness by treatment group, age or gender

```{r}
data_expanded %>% 
  mutate(missing_mem_comp = case_when(
    is.na(mem_comp) ~ 1,
    TRUE ~ 0
  )) %>% 
  select(-subject_id, -mem_comp) %>% 
  tbl_summary(by = missing_mem_comp) %>% 
  add_p()
```

Distribution of the mem_comp variable

Sort of skewed, but not too badly

```{r}
ggplot(data_expanded, aes(x = mem_comp)) + 
  geom_histogram()
```


# Impute data

Assume that the data are missing at random, conditional on the data available (age, gender, days)

Take clustering into account during imputation

```{r message=FALSE}
# predictive mean matching
# semi-parametric imputation approach
# https://stats.idre.ucla.edu/r/faq/how-do-i-perform-multiple-imputation-using-predictive-mean-matching-in-r/
# for each missing value, it fills in a value randomly from among the a observed donor values from an observation whose regression-predicted values are closest to the regression-predicted value for the missing value from the simulated regression model (Heitjan and Little 1991; Schenker and Taylor 1996)
imputed_df_pmm <- mice(data_expanded, method = "pmm", m = 5, seed = 1123)

# imputation method used for each variable
# only mem_comp is imputed
imputed_df_pmm$method
summary(imputed_df_pmm)

# predictor matrix
imputed_df_pmm$predictorMatrix

# default plots output
plot(imputed_df_pmm)

# here is the imputed data
imputed_df_pmm$imp$mem_comp

# get back the dataset with imputed data
data_pmm <- complete(imputed_df_pmm)

# only 188
nrow(data_pmm)

# fit model on imputed data
model_pmm <- with(data_pmm, lmer(mem_comp ~ treatment_group + day + treatment_group*day + age + gender + (1 | subject_id), data = data_pmm))

# table of model results on imputed data
# interaction between tx * day is not significant -> remove and re-run
model_pmm %>% 
  tbl_regression() %>% 
  add_global_p()

# w/o interaction term
model_pmm_v2 <- with(data_pmm, lmer(mem_comp ~ treatment_group + day + age + gender + (1 | subject_id), data = data_pmm)) 

# table of results
# nothing statistically significant
model_pmm_v2 %>% 
  tbl_regression() %>% 
  add_global_p()

# doesn't work
# summary(pool(model_pmm))
```

Spaghetti plot with imputed data

```{r}
data_pmm %>% 
  mutate(measurement_time = case_when(
    day == 0 ~ 1,
    day == 5 ~ 2,
    day == 19 ~ 3,
    day == 90 ~ 4
  )) %>% 
  ggplot(aes(x = day, y = mem_comp)) +
  geom_point(aes(color = treatment_group, group = subject_id)) +
  geom_line(aes(color = treatment_group, group = subject_id)) +
  geom_smooth(color = "gray", size = 2, method = "loess", se = FALSE) +
  labs(x = "Day",
       color = "Treatment",
       y = "Cognitive Rating",
       caption = "Higher cognitive rating indicates better memory.") +
  facet_grid(.~treatment_group)
```
